\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-11-16 paper/main.tex

% ----------------------------------------------- ILLUSTRATIVE EXAMPLES ------------------------------------------------
\section{Illustrative Examples}
% Provide at least one illustrative example to demonstrate the major functions.

@@

Our work also has several more subtle ``advanced'' use cases and
potential impacts. Whereas Python's built-in \texttt{import}
statement is agnostic to packages' version information, \texttt{smuggle}
statements (when combined with onion comments) are version-sensitive.
And because onion comments are parsed at runtime, required packages and
their specified versions are installed in a just-in-time manner. Thus, it is possible
in most cases to \texttt{smuggle} a specific package version or revision even if
a different version has already been loaded. This enables more complex uses
that take advantage of multiple versions of a package within a single interpreter
session. This could be useful in cases where specific features are added or
removed from a package across different versions, or in comparing the
performance or functionality of particular features across different versions of
the same package.

A second advanced use case is in providing a proof-of-concept of how
one can add new ``keyword-like'' operators to the Python language by
leveraging notebooks' error-handling mechanisms. This could lead to
exciting new tools that, like \texttt{davos}, extend the Python
language in useful ways within notebook-based environments. We note that
our approach to adding the \texttt{smuggle} keyword to Python when
\texttt{davos} is imported into a notebook-based environment also has
the potential to be exploited for more nefarious purposes. For
example, a malicious user could use a similar approach (e.g., in a
different library) to substantially change a notebook's functionality
by adding new \textit{unexpected} keyword-like objects (e.g., based
around common typos). This could lead to difficult-to-predict changes
in a notebook's behavior once the malicious library was imported.
This highlights an important reason why security-conscious users would
be well-served to only make use of libraries from trusted sources, or
whose code is publicly available for review.

>>>>

% ----------------------------------------------- ILLUSTRATIVE EXAMPLES ------------------------------------------------
\section{Illustrative Example}\label{sec:illustrative-example}
% Provide at least one illustrative example to demonstrate the major functions.

@@

Our work also has several more subtle ``advanced'' use cases and
potential impacts. Whereas Python's built-in \texttt{import}
statement is agnostic to packages' version information, \texttt{smuggle}
statements (when combined with onion comments) are version-sensitive.
And because onion comments are parsed at runtime, required packages and
their specified versions are installed in a just-in-time manner. Thus, it is possible
in most cases to \texttt{smuggle} a specific package version or revision even if
a different version has already been loaded. This enables more complex uses
that take advantage of multiple versions of a package within a single interpreter
session (e.g., see Section~\ref{sec:illustrative-example}). This could be useful in cases where specific features are added or
removed from a package across different versions, or in comparing the
performance or functionality of particular features across different versions of
the same package.

A second more subtle impact of our work is in providing a proof-of-concept of how the ability to add new ``keyword-like'' operators to the Python language could be specifically useful to researchers. With \texttt{davos}, we accomplish this by leveraging IPython notebooks' internal code parsing and execution machinery. We note that, while other popular packages similarly use these mechanisms to providing notebook-specific functionality (e.g., \cite{Hunt07,HeusEtal18a}), this approach also has the potential to be exploited for more nefarious purposes. For example, a malicious user could design a Python library that, when imported, substantially changes the notebook's functionality by adding new \textit{unexpected} keyword-like objects (e.g., based around common typos). We also note that this implementation approach means \texttt{davos}'s functionality is currently restricted to IPython notebook environments, as would be that of any potential similar tools that enable user-defined keywords. However, there has been early-stage discussion of providing this sort of syntactic customizability as a core feature of the Python language, including a draft proposal \cite{Shan20}. In addition to enabling \texttt{davos} to be extended for use outside of notebooks, this could lead to exciting new tools that, like \texttt{davos}, extend the Python language in useful and more secure ways.
